\begin{acknowledgements}
\phantomsection % required for using hyperref
\addcontentsline{toc}{chapter}{Acknowledgements}

When I first started this research I was in late 20's. Now I am over 40
years old. After spending almost thirteen years of my life to complete this
thesis, I have now learned one thing –- I could never have done any of this,
particularly the research and writing that went into this thesis, without God
who helped me through many people around me; supervisors, family and friends.

\vspace{3 mm}

First, I would like to thank my three supervisors, David Favis-Mortlock, John
Boardman and Richard Washington for their constant supports throughout my
research. David, I thank you specially for your encouragements, supports and
countless comments on my thesis. Even I would not enjoy reading my writings for
that many times. Still you gave me all those detailed comments.

\vspace{3 mm}

I have been so fortunate to have my parents and parents-in-law who have been
supporting my study in various ways for many years. Without them, it would
not have been possible to start, continue and complete the degree in Oxford. My
deepest appreciations goes to them.

\vspace{3 mm}

I would also like to thank Tony Lemon, my college advisor, and Brenda Boardman
for their kind and thoughtful words.

\vspace{3 mm}

It was my great pleasure to be a part of Mansfield College, Environmental Change
Institute, School of Geography and the Environment and University of Oxford. I
will be greatly missing the research environment and academic supports that I
took for granted.

\vspace{3 mm}

My sincere appreciation extents to Wonsuk and Julie Ma who became my parent
figures in Oxford. Thank you so much for your unconditional love and kindness.

\vspace{3 mm}

There also have been many friends who prayed for me and my study: David and
Safiya Nyirongo, Mark and Asja Powys-Smith, Mike and Jacque Morrison, Hugo and
Parveen Marques, Mezza and Roz, Thomas, Marco-Antonio, Bill, Gordon Hickson, Bob
and Jenny Oh, Choongkook and Sewon, Kongju Lee, Alastair and Juliet Curry, Judy
and Lily's close friends: Lidia, Shanon and Soonah. I thank you all for your
prayers.

\vspace{3 mm}

Finally, I would like to dedicate this work to my wife, Lily, for her patience
and love during so many lonely days and late nights. I do not know how I can
repay you all those days. I thank you and love you.

\end{acknowledgements}