\chapter{EUROSEM Input Data}
\label{sec:EUROSEMInputData}

\section{SYSTEMS}
\label{sec:SYSTEMS}

\footnotesize

\paragraph{NELE [1]}
This defines the total number of elements in the catchment. Its value should be
the same as the number of elements entered under ELEMENT NUM. (J) in the
Rainfall Data File (XXX.pcp). Since, in this example, only one slope plane
(Woodingdean D) was being considered, we entered NELE = 1. Woodingdean site is
divided into 9 slopes (D ~ L).
 
\paragraph{NPART [0]}
This relates to a component of the KINEROS model which describes the settling
of sediment in ponds. It is not used in the present version of EUROSEM. A value
of 0 should always be set here.
 
\paragraph{CLEN (m) [135]}
This is the characteristic length of overland flow and represents the longest
possible length in a series of cascading planes or channels. Since the example
erosion plot was being treated as one slope element, CLEN was set here as equal
to the downslope length of the plot (the maximum lengths of longest channel),
i.e. 135 m.
 
\paragraph{TFIN (min) [870]}
This is the total computational time (min) for which the model is to be run.
Its value must be less than the end-time of the last time-depth pair in the
Rainfall Data File. The value of TFIN will depend upon the duration of the storm
and the response time of the catchment. It should be sufficient to contain the
hydrograph of surface runoff and should therefore extend from the start of the
rainfall to the time that surface runoff on the hillslopes ceases. For the storm
considered here, the last time-depth pair ends at 900 minutes, so we set TFIN =
870 min. Actual rainfall duration for the storm is 870 minutes. 30 minutes were
added to indicate the end of storm (no rain for 30 min).
 
\paragraph{DELT (min) [1]}
This defines the time increment used in the simulations. Ideally, this should
be as short as possible. However, the total number of time steps, defined as
TFIN/DELT should not exceed 1000 in which case the model will pause and a
warning message will appear. For this example, I choose a value of DELT = 1.0.
(870/1 = 870 $<$ 1000)
 
\paragraph{THETA [0.7]}
This is a weighting factor used in the finite difference equations in KINEROS
for routing overland flow and channel flow. It should have a value between 0.5
and 1.0. A value of 0.7 is recommended and this was the value we chose.
 
\paragraph{TEMP [2.55]}
The air temperature (\textdegree\ Celsius) at the start of the storm should be
set here. It is used in the model to compute the kinematic viscosity of water.
Mean temperature on the day of storm was chosen. (Max = 4.8 \textdegree\ C and
Min = 0.3 \textdegree\ C )
 
\section{OPTIONS}
\label{sec:OPTIONS}

No changes should be made to the entries under this heading.
EUROSEM is designed to operate with values of 2 under both entries.
 
\paragraph{NTIME [2]}
This is the code for the time units used in the model. NTIME = 1 for seconds
and NTIME = 2 for minutes. The value of 2 should always be used with the present
version of EUROSEM.
 
\paragraph{NEROS [2]}
This allows the user to call or reject the erosion option in the model. With
values set at 0 or 1, the erosion option is not called and only the hydrological
calculations are made. A value of 2 calls the erosion option which, in this case
is EUROSEM.
 
\section{COMPUTATION ORDER}
\label{sec:COMPUTATIONORDER}

This heading describes the order in which the plane and channel elements
comprising the catchment must be organised to provide the correct cascading
sequence for the movement of runoff and sediment downslope and downstream.
 
\paragraph{COMP. ORDER (NLOG) [1]}
This denotes the order of calculation. Each entry must therefore be in
numerical sequence.
 
\paragraph{ELEMENT NUM. (J) [1]}
This defines the corresponding element number for each entry in the sequence.
The element numbers need not be in numerical order. The total number of elements
listed here should be the same as the total number entered under ELE. NUM. (J)
in the Rainfall Data File and correspond to the number entered under NELE above.
Since only one slope plane was being considered at Woburn, NLOG was set to 1 and
ELEMENT NUM. was therefore also equal to 1.
 
\section{ELEMENT WISE INFO}
\label{sec:ELEMENTWISEINFO}

This heading gives the data on the catchment characteristics of each element.
The number by which each element is known must be the same as that listed above
under ELEMENT NUM, where the computational order is defined, and also under ELE.
NUM. (J) in the Rainfall Data File.
 
\paragraph{J [1]}
This represents the number of the element. J = 1 for the first element, J = 2
for the second element, and so on. In the example being used here, there was
only one element, so J = 1.
 
\paragraph{NU [0]}
This denotes the number of the element which contributes runoff and sediment to
the upslope boundary. Since there was only one element, there were no upslope
contributing elements, so NU = 0.
 
\paragraph{NR [0]}
This entry applies to elements which are channels and denotes the number of the
hillslope elements contributing flow to the channel from the right-hand side
when viewed in the direction of flow, i.e. facing downstream. For hillslope
elements, as here, NR = 0.
 
\paragraph{NL [0]}
This entry similarly applies to channels and denotes the number of the
hillslope element contributing flow to the channel from the left-hand side. For
hillslope elements, as here, NL = 0.
 
\paragraph{NC1 [0]}
This entry also applies to channels and denotes the number of the first channel
element contributing flow to the channel from upstream. For hillslope elements,
NC1 = 0.
 
\paragraph{NC2 [0]}
This entry denotes the number of the second channel element contributing flow
to the channel from upstream. It is relevant for channels downstream of a
confluence so that there are two contributing channel elements at the upstream
end. For hillslope elements, NC2 = 0.
 
\paragraph{NPRINT [1]}
This controls the amount of information provided in the auxilary output file.
In our case it is set to 1.
 
\paragraph{XL (m) [135.0]}
This is the length of the element (in meter). Since the erosion plot was 135 m
long, XL = 135.0.
 
\paragraph{W (m) [50]}
This is the width of the element (m). Since the erosion plot was 50 m wide, W =
50.0. It should be noted that W = 0.0 if the element being described is a
channel.
 
\paragraph{S [0.14]}
This is the average slope of any rills on the element (m/m), measured in the
direction of maximum slope, i.e. at right angles to the contour. Since the
average slope of the plot was measured in the field at 14 per cent, we entered S
= 0.14.
 
\paragraph{ZR [0]}
This is the side slope of the right-hand side of the channel, assuming a
trapezoidal shape and expressing the slope as 1:ZR. Since we were dealing with a
plane element, there were no channels, so ZR = 0.
 
\paragraph{ZL [0]}
This is the side slope of the left-hand side of the channel, assuming a
trapezoidal shape and expressing the slope as 1:ZL. Since we were dealing with a
plane element, there were no channels, so ZL = 0.
 
\paragraph{BW [0]}
This is the bottom width (m) of the channel, assuming a trapezoidal shape.
Since we were dealing with a plane element, there were no channels, so BW = 0.0.
 
\paragraph{MANN (Rill) [0.125]}
This is the value of Manning's n for the rill channels (concentrated flow
paths) on the element, taking account of the combined effects of soil particle
roughness, surface microtopography and plant cover on the element. For the sandy
loam soil in a smooth seed-bed, a typical value would be n = 0.015. For wheat, n
ranges from 0.01 to 0.30, depending on the percentage cover and planting
density. For the smooth seedbed and 10 per cent cover prevailing at the time of
the storm, we estimated a value at the lower end of the range, e.g. 0.04.
The value for Manning's n should be further adjusted to take account of rock
fragments or stones in the surface soil, using equation A3.2 in appendix 3.
 
\paragraph{Mann (IR) [0.170]}
This is the value of Manning's n for the interrill area of the element, again
taking into account soil particle roughness, surface microtopography and plant
cover. For the smooth surface and cover of the element in question, the same
value was chosen as for Manning's n in the rills. We therefore entered IRMANN =
0.04.
As with the case above, the Manning's n value should be adjusted, if necessary,
for rock fragments or stones in the surface soil, using equation A3.2 in
appendix 3. No such adjustment was needed for Woburn. You should note that the
model will further adjust the value of IRMANN to take account of the level of
roughness on the interrill area, as expressed by the downslope roughness ratio,
RFR (Appendix 6).
 
\paragraph{FMIN (mm/h) [3.0]}
This is the saturated hydraulic conductivity of the soil (mm/h). This should be
the value for the soil itself and should not be adjusted for plant cover or
stoniness. These adjustments are made within the model itself, as functions of
input data on PBASE and ROC respectively. If FMIN has been measured for soils
with a vegetation or stone cover, the measured value should be used. The input
values for PBASE and ROC should then be set to zero so that no further
adjustment is made to the FMIN value within the model.
Effective hydraulic conductivity of the soil should be 3 mm/h (from WEPP
output).
 
\paragraph{G (mm) [480.0]}
This is the effective net capillary drive of the soil (mm), as described in
Section 3.3.1. From Table A4.1, a value of 480 was chosen for a silt loam soil,
so here G = 480.
 
\paragraph{POR [0.5]}
This is the porosity of the soil (\% v/v). From Table A4.1, a value of 0.50 was
chosen for a silt loam soil and we entered POR = 0.50.
 
\paragraph{ThI [0.4]}
This is the volumetric moisture content of the soil at the start of the storm.
This has to be estimated in relation to the time since it last rained and the
speed with which the soil dries out. As explained in Appendix 4, THI will take a
value between the maximum moisture content of the soil (THMX) and the moisture
content at wilting point. Since the storm occurred in the middle of a wet spell
of weather, the soil had had little opportunity to dry out between storms. A
rather high value of THI = 0.4 was therefore chosen.

\paragraph{ThMX [0.42]}
This is the maximum moisture content of the soil. From Table A4.1, we chose a
value of THMX = 0.42.
 
\paragraph{ROC [0.0]}
This is the proportion (\% v/v) of the soil occupied by stones and rocks. Since
the silt loam soil at Woodingdean is very stony, we could have entered ROC =
0.381 (from the WEPP input, andover.sol). However, a value of ROC = 0.0 was used
as the input value for FMIN is assumed as a measured one, which already takes
account of the presence of rock fragments or stones.
 
\paragraph{RECS (mm) [1.0]}
This is the infiltration recession factor and is defined as the average maximum
local difference in microrelief (mm). Based on field measurements of surface
roughness (Appendix 6), a value of RECS = 20.0 was selected.
It should be noted that a value of RECS $>$ 0 must always be entered.
 
\paragraph{DINT (mm) [3.0]}
This is the maximum interception storage of the plant cover (mm). From Table
A7.1, for winter-sown wheat, a value of DINTR = 3.0 was chosen.
 
\paragraph{DEPNO [15.0]}
This denotes the average number of rills (concentrated flow paths) across the
width of the slope plane. Since the erosion plot is ploughed up and down slope,
the plough furrows act as concentrated flow paths. Based on field observations,
an average of ten paths was recorded, using the procedure shown in Appendix 8. A
value of DEPNO = 15.0 was therefore entered.
 
\paragraph{RILLW (m) [0.02]}
This is the average bottom width (m) of a concentrated flow path or rill. A
arbitrary value of RILLW = 0.02 was entered.
A flat surface would be assigned a value of RILLW = 0.0.
 
\paragraph{RILLD (m) [0.03]}
This is the average depth (m) of a concentrated flow path or rill. A arbitrary
value of RILLD = 0.03 was entered.
 
\paragraph{ZLR [10.0]}
This denotes the average side slope of a concentrated flow path (rill),
expressed as 1:ZLR.  A arbitrary value of ZLR = 10.0 was entered.
 
\paragraph{RS [1.0]}
If RS = 0, the model assumes that the values of RILLW and RILLD entered above
apply for the whole length of the element. If RS = 1, the model assumes the
values apply to the rill at the lower end of the element and scales the values
to smaller dimensions with distance upslope. In this case, the scaling option
was selected, so we entered RS = 1.
 
\paragraph{RFR [1.0]}
This is the downslope roughness ratio. Based on field measurements, using the
procedure described in Appendix 6 and illustrated in Figure A6.1, a value of RFR
= 1.0 was obtained and entered.
Although this value is much lower than those listed in Table A6.1, it is a
typical value for a relatively smooth surface. As stated earlier when choosing a
value for Manning's n, the condition of the ground at the time of the storm was
a smooth seed-bed flattened by several months of raindrop impact.
 
\paragraph{SIR (m/m) [0.2]}
This is the interrill slope. For unrilled plane elements, this is the average
slope of the plane. For channel elements, this is the average slope of channel.
For a plane element with rills, SIR is defined as the average ground slope
followed by overland flow as it passes over the interrill area into the rills
(see Appendix 2). The average slope of the rills should be entered under S. A
arbitrary value of SIR = 0.5 was entered.
 
 
\paragraph{COVER [0.3]}
This is the effective percentage canopy cover of the vegetation. Strictly it
refers to the proportion (between 0 and 1) of the ground surface obscured by the
vegetation when viewed vertically from above. The value should take account of
ground vegetation, mulches and any litter layer as well as trees and bushes.
Since, at the time of the storm, this was estimated at 30 per cent, a value of
COVER = 0.3 was entered.
 
\paragraph{SHAPE [1]}
This refers to the shape of the leaves. SHAPE = 1 for bladed leaves and needle
leaves. SHAPE = 2 for broad leaves. Since the crop was wheat, we entered SHAPE =
1. Conceptually, the SHAPE factor describes, in a simplified way, the
relationship between the size of the leaves and the median volume drop diameter
of the rainfall. A value of 0, to be entered when there is no vegetation cover,
will cause stemflow to be set zero.

\paragraph{PLANGLE (degree) [85.0]}
This is the average acute angle (degrees) between the plant stems and the
ground surface. Based on the model manual (Table A7.2), a mean value of PLANGLE
= 85\textdegree\ was entered (wheat = 80--90).
 
\paragraph{PLANTBASE [0.0]}
This is the percentage basal area of the vegetation cover. From Table A7.3, we
can see that the value for small grains (wheat, barley, rice) ranges from 0.2 to
0.3, depending on the planting density. As this may be assumed to be high, a
value of 0.3 is chosen. Since the percentage plant cover was only 30 per cent,
the value was reduced accordingly and we could have entered PLANTBASE = 0.09.
However, it should be noted that if the value entered for FMIN has been
determined in the field for vegetated conditions, PLANTBASE should be set to
0.0. This avoids further adjustment of the FMIN value within the model to allow
for the effect of the vegetation cover.
 
\paragraph{PLANTH (cm) [100]}
This is the average height of the plant canopy (cm) above the ground surface.
From Table A7.2, wheat has a height of 50 -- 150 cm. Since, the purpose is to
describe the fall height of intercepted raindrops, any ground vegetation,
mulches and litter layer should be considered. A value of 100 (cm) was entered
for this example.
 
\paragraph{DERO (m) [2.0]}
This is the maximum depth (m) to which erosion can proceed before a resistant
or non-erodible layer (e.g. plough pan or concretionary horizon) below the soil
surface. Once erosion reaches this depth, the model prevents further downcutting
by rills; from then on the rills are only able to erode by widening their
channels. Since there are layers with high clay contents from 2 m, we entered
DERO = 2.0.
 
\paragraph{ISTONE [+]}
An indicator of the effect of rock fragments on the surface of the soil on the
saturated hydraulic conductivity (see Appendix 5).
A value of +1 should be used where the rock fragments sit on the surface and
protect the soil from structural breakdown due to raindrop impact; or where the
rocks either sit on or are fully embedded in a soil with high macroporosity,
e.g. due to recent tillage. In this instance, the rock fragments will enhance
infiltration.
A value of -1 should be used where the rocks are partially embedded within or
sit on top of a sealed surface which will reduce infiltration.
When PAVE value is set to zero, this value is not considered in the model.
 
\paragraph{D50 ($\mu$m) [35.0]}
This is the median particle size of the soil as obtained from standard
particle-size analysis, using the USDA system to define textural classes (i.e.
clay: $<$ 0.002 mm; silt: 0.002 - 0.05 mm; sand 0.05 - 2.00 mm).
A value of D50 = 35 was entered.
 
\paragraph{EROD (g/J) [0.8]}
This is the detachability of the soil particles by raindrop impact (g/J). From
Table A9.1, for a silt loam soil, a minimum value of EROD = 0.8 was selected.
 
\paragraph{SPLTEX [2.0]}
This is the value of the exponent relating detachment of soil particles by
raindrop impact to the depth of water on the soil surface. The current version
of EUROSEM uses a constant value of 2.0 for this exponent.
 
\paragraph{COH (kPa) [6.3]}
This is the cohesion of the soil as measured in the field with a torvane (Soil
Test CL-600) after the soil has been saturated (see Appendix 9). Guide values
for soils with different textures are given in Appendix 9. The value should be
adjusted to take account of the effects of the root system of the vegetation.
From field measurements on the bare saturated silt loam soil (compacted),
cohesion is low at about 6.0 kPa. From Table A9.3, assuming that wheat has a
similar effect to barley, an increase in cohesion of between 0.2 and 0.6 kPa may
be expected as a result of root reinforcement. For a crop at the stage of 30 \%
cover, we might estimate an increase at the lower end of the range, say 0.3 kPa.
If this is added to the cohesion value for the bare soil, we get a total
cohesion of 6.0 + 0.3 kPa = 6.3 kPa. A value of COH = 6.3 was therefore entered.
 
\paragraph{RHOS (Mg/m3) [2.65]}
This is the specific gravity of the sediment particles. This is normally set at
2.65 Mg/m3.
 
\paragraph{PAVE [0.0]}
This is the fraction of the surface occupied by non-erodible material (e.g.
rock fragments, concrete, tarmac). It is used in EUROSEM to reduce the rate of
soil detachment by raindrop impact in direct proportion to the area occupied by
non-erodible surfaces and also to influence the way rock fragments affect the
saturated hydraulic conductivity of the soil (see ISTONE). This value was set to
zero to prevent any further adjustment of FMIN.
 
\paragraph{SIGMAS [0]}
This is the standard deviation of the sediment particle diameter ($\mu$m) for
any element immediately upslope of a pond. It is used within KINEROS for
modelling the process of sedimentation in a pond or reservoir. Since current
version of EUROSEM does not deal with ponds, SIGMAS was set = 0.0.
 
\paragraph{MCODE [0]}
The value chosen for MCODE allows the user to choose the equations used in
EUROSEM to simulate sediment transport by interrill flow.
MCODE = 1 selects the equations proposed by Everaert (1992) which relate
specifically to interrill flow. MCODE = 0 selects the equations proposed by
Govers (1990) for rill flow and applies them to both interrill and rill flow.
See the model documentations for more details.
